### 2017-05-22 abstract.tex
[draft] abstract and structure of paper added

We study the quantum quench problem for a class of bosonic conformal interface by computing the fidelity and Loschmidt echo. The scattering matrix $S(\theta)$ associated with the permeable interfaces are classified into two sectors of $O(2)$ group. The fidelity and Loschmidt echo decay algebraically as $t^{-\alpha}$ after the quench. The exponent $\alpha$ is found to depend on the change of $\theta$ as a quadratic relation if the boundary conditions (before and after quench) are in the same sectors of $O(2)$, while remain $\frac{1}{4}$ if they belong to different sectors. We confirm our results by various numerical simulations. The winding modes of compact bosons will have no contribution at leading order. We expect such conformal interface can be realized in a wide range of experiments of one-dimensional bosonized electron gas.

%1. cut and joint two different CFT

%7 Experiment