### 2017-05-23 compact_boson.tex

In this appendix, we address the issue of winding modes of compact bosons. In the main text, we have exclusively worked with the oscillator modes of free bosons. Here, we shall show that winding modes for compactified bosons will have no contribution to the fidelity or Loschmidt echo in the leading order. Therefore, our results are ready to be applied in the case where two compactified bosons of different radii are connected by a conformal interface\cite{PhysRevLett.118.136801}. \todo[inline]{say intuition why not matter, mention application to Laughlin, etc.}

Our derivation follows the general multi-component Boson constraints in Ref.~\onlinecite{affleck_quantum_2001,oshikawa_boundary_2010}. A review of the detailed parameterization of the states can be found in Ref.~\onlinecite{sakai_entanglement_2008}.

\subsection{Mode Expansion of Compact Boson}
Suppose the boson is compactified as $\phi =  \phi + 2\pi R$, then mode expansion will have zero mode

\begin{widetext}
\begin{equation}
\begin{aligned}
\phi( x, t ) &= \phi_0 - \frac{2\pi m R}{L}x + \frac{\pi_0}{gL} t  + \frac{i}{\sqrt{4 \pi g}}\sum_{n\neq0}\frac{1}{n}(a_ne^{2\pi i n(x-t)/L}+ \bar{a}_{n} e^{2\pi i n(x+t)/L})\\  
&= \phi_0 - \frac{2\pi m R}{L}x + \frac{n}{gLR} t  +  \frac{i}{\sqrt{4 \pi g}}\sum_{n\neq0}\frac{1}{n}(a_ne^{2\pi i n(x-t)/L}+ \bar{a}_{n} e^{2\pi i n(x+t)/L})\\ 
\end{aligned}
\end{equation}
where the conjugate momentum $\pi_0$ takes discrete eigenvalues $\frac{n}{R}$. In the holomorphic coordinates
\begin{equation}
\label{eq:zzbar}
z = \exp( 2 \pi i \frac{t - x}{L}) \qquad \bar{z} = \exp( 2 \pi i \frac{t + x}{L})
\end{equation}
the mode expansion becomes
\begin{equation}
\label{eq:boson-mode-exp}
\begin{aligned}
\phi( z, \bar{z}) = &\phi_0 -i \left( \frac{n}{4\pi g  R} + \frac{m R }{2} \right)  \ln z + \frac{i}{\sqrt{4\pi g}} \sum_{n\ne 0 } \frac{a_n}{n} z^{-n } -i \left( \frac{n}{4\pi g R} - \frac{m R }{2} \right)  \ln \bar{z} + \frac{i}{\sqrt{4\pi g}} \sum_{n\ne 0 } \frac{\bar{a}_n}{n} \bar{z}^{-n } 
\end{aligned}
\end{equation}
\end{widetext}
We further identify 
\begin{equation}
\begin{aligned}
  a_0 &= \sqrt{ 4 \pi g } \left( \frac{n}{4\pi g R} + \frac{m R }{2} \right) \\
   \bar{a}_0 &= \sqrt{ 4 \pi g } \left( \frac{n}{4\pi g R} - \frac{m R }{2} \right)
  \end{aligned}
\end{equation}
and then 
\begin{equation}
\begin{aligned}
i \partial_z \phi =  \sum_n \frac{1}{\sqrt{4\pi g}} a_n z^{-n-1} 
\end{aligned}
\end{equation}

\subsection{Gluing Condition for Winding Modes}
\label{app_sub:compact_gluing_boundary}

Following the treatment of boundary CFT in Ref.~\onlinecite{di_francesco_conformal_1997}, we shall quantize the theory by imposing \emph{equal-time} commutator for $a_n$ and $\bar{a}_n$ in Eq.~\eqref{eq:boson-mode-exp}. The conformal boundary state will be defined on a constant time slice, say $t=0$. For the permeable wall connecting two compactified bosons, this results in folding $t\rightarrow-t$ rather than $x\rightarrow-x$ as in the main text. Since the two cases are {\bf T}-dual to each other, the results are not affected.

%Before we invoke the folding trick, we need to note one important caveat here. In the usual treatment of boundary CFT, for example in , the field is quantized on the real axis. As emphasized in Sec.~\ref{sec:bosonic_conformal_interface}, the gluing condition becomes an operator identity of creation and annihilation operators, which is valid \emph{only} when acting on the conformal boundary states.

We recall that the gluing condition is written as
\begin{equation}
\begin{aligned}
\label{eq:def_S_in_app}
\begin{pmatrix}
\partial_+\phi^1\\
\partial_-\phi^2
\end{pmatrix}
=S(\theta)
\begin{pmatrix}
\partial_-\phi^1\\
\partial_+\phi^2
\end{pmatrix}
\end{aligned}
\end{equation}
The folding sends $\phi^2(t)\rightarrow\phi^2(-t)$ for $t<0$, and the $\partial_t \phi^2$ term changes sign, we therefore have 
\begin{equation}
\begin{aligned}
\label{eq:def_S_in_app_2}
\begin{pmatrix}
\partial_+\phi^1\\
-\partial_+\phi^2
\end{pmatrix}
=S(\theta)
\begin{pmatrix}
\partial_-\phi^1\\
-\partial_-\phi^2
\end{pmatrix}
\end{aligned}
\end{equation}
For convenience, we shift $\phi^2\rightarrow-\phi^2$ by a sign. This amounts to reparameterize the S-matrix and exchange the two sectors of $O(2)$ group. But this will not change our conclusion in the main text. 
\todo[inline]{Check the parameterization} 
Upon mode expanding $\phi^{1,2}$ in the holomorphic coordinates as in Eq.~\eqref{eq:boson-mode-exp}, we have
\begin{equation}
\begin{aligned}
\label{eq:def_S_in_app_2}
\sum_{n\in\mathbb{Z}}
\begin{pmatrix}
\bar{a}_n^1\bar{z}^{-n}\\
\bar{a}_n^2\bar{z}^{-n}
\end{pmatrix}
=S(\theta)
\sum_{n\in\mathbb{Z}}
\begin{pmatrix}
a_n^1{z}^{-n}\\
a_n^2{z}^{-n}
\end{pmatrix}
\end{aligned}
\end{equation}
As explained above, the boundary condition is imposed at $t=0$, where $\bar{z}=z^{-1}$, therefore we have
\begin{equation}\begin{aligned}
a^i_n-(S^{-1}_{ij})\bar{a}^j_{-n}=0
\end{aligned}\end{equation}
The solution of the $n \ne 0$ constraints is exactly the boundary state in Eq.~\eqref{eq:bd_state}.

%We recall that the gluing condition can written as
%\begin{equation}\begin{aligned}
%\label{eq:def_M}
%\begin{pmatrix}
%\partial_t\phi^1\\
%\partial_x\phi^1
%\end{pmatrix}_{t=0^-}
%=M\begin{pmatrix}
%\partial_t\phi^2\\
%\partial_x\phi^2
%\end{pmatrix}_{t=0^+}
%\end{aligned}\end{equation}
%where the derivatives are evaluated at the appropriate limit. In this appendix, we shall take the definition $x^\pm=x\pm t$ and $\partial_\pm=\partial_x\pm\partial_t$. Therefore we establish the scattering matrix as
%\begin{equation}
%\begin{aligned}
%\label{eq:def_S_in_app}
%\begin{pmatrix}
%\partial_+\phi^1\\
%\partial_-\phi^2
%\end{pmatrix}
%=S(\theta)
%\begin{pmatrix}
%\partial_-\phi^1\\
%\partial_+\phi^2
%\end{pmatrix}
%\end{aligned}
%\end{equation}
%
%The folding sends $\phi^2(t)\rightarrow\phi^2(-t)$ for $t<0$, and the $\partial_t \phi^2$ term changes sign, we therefore have 
%\begin{equation}
%\begin{aligned}
%\label{eq:def_S_in_app_2}
%\begin{pmatrix}
%\partial_+\phi^1\\
%\partial_+\phi^2
%\end{pmatrix}
%=S(\theta)
%\begin{pmatrix}
%\partial_-\phi^1\\
%\partial_-\phi^2
%\end{pmatrix}
%\end{aligned}
%\end{equation}
%We note that the parameterization of S-matrix is \emph{different} from Eq.~\eqref{eq:def_S}, but this will not change our conclusions in the main text. Upon mode expanding $\phi^{1,2}$ in the holomorphic coordinates as in Eq.~\eqref{eq:boson-mode-exp}, we have
%\begin{equation}
%\begin{aligned}
%\label{eq:def_S_in_app_2}
%\sum_{n\in\mathbb{Z}}
%\begin{pmatrix}
%\bar{a}_n^1\bar{z}^{-n}\\
%\bar{a}_n^2\bar{z}^{-n}
%\end{pmatrix}
%=-S(\theta)
%\sum_{n\in\mathbb{Z}}
%\begin{pmatrix}
%a_n^1{z}^{-n}\\
%a_n^2{z}^{-n}
%\end{pmatrix}
%\end{aligned}
%\end{equation}
%\todo[inline]{Check the sign and parameterization of S}
%As explained above, the boundary condition is imposed at $t=0$, where $\bar{z}=z^{-1}$, therefore we have
%\begin{equation}\begin{aligned}
%a^i_n-(-S^{-1}_{ij})\bar{a}^j_{-n}=0
%\end{aligned}\end{equation}
%The solution of the $n \neq 0$ constraints is exactly the boundary state in Eq.~\eqref{eq:bd_state}, with an extra minus sign. Following our boundary state calculation in the main text, it is clear that this extra minus sign, which is the result of {\bf T}-duality, does not affect our conclusion.

We specialize to $S = S_1$ to solve the $ n = 0$ constraint. We introduce the compactification lattice and its dual\cite{affleck_quantum_2001,oshikawa_boundary_2010}
\begin{equation}
\label{eq:lattice}
\vec{M} = (m_1 2 \pi R_1, m_2 2\pi  R_2)^\top, \quad  \vec{M}^* = (\frac{n_1}{R_1}, \frac{n_2}{R_2})^\top
\end{equation}
to rewrite the zero mode part as 
\begin{equation}
  a_0^i - S^{-1} _{ij} \bar{a}_{0}^j = 0 \quad \implies \quad ( \vec{M} + \frac{1}{g}\vec{M}^* ) = S_1 ( -\vec{M} + \frac{1}{g}\vec{M}^* )
\end{equation}
which is basically the multi-component boson winding constraints given in \onlinecite{affleck_quantum_2001,oshikawa_boundary_2010}. The solution gives the interface parameter $\lambda$
\begin{equation}
\begin{aligned}
\label{eq:S_1_constraint}
\lambda = \tan\theta=\frac{n_2R_1}{n_1R_2}=-\frac{m_1R_1}{m_2R_2}
\end{aligned}
\end{equation}
and the conformal boundary state reads
\begin{equation}\begin{aligned}
g_{S_1}\sum_{S_1}e^{in_1\phi_0-im_1\bar{\phi}_0}|n_1,m_1\rangle|n_2,m_2\rangle
\end{aligned}\end{equation}
where $\sum_{S_1}$ is the summation consistent with the constraint in Eq.~\eqref{eq:S_1_constraint}. The g-factor can only be determined by Cardy condition\cite{cardy_boundary_2004}. Since it is not important for what follows, we shall not repeat the calculation here. Since $S = S_2$ is effectively $S_1$ on the dual boson, we can expect that it will end up in a similar parameter $\lambda$ and boundary state. We omit the expression here because the main text does not use it.

\subsection{Winding Mode Contribution to the Partition Function}
\label{app_sub:winding_contribution}
We now calculate the winding mode part of the partition function as shown in Fig.~\ref{fig:fig_lambda_1_lambda_2}
\begin{equation}\begin{aligned}
Z=\langle\lambda_2|e^{-\pi H}|\lambda_1\rangle\qquad H=\frac{2\pi}{\beta}(L_0+\bar{L}_0)
\end{aligned}\end{equation}
To that end, we can simply replace $L_0$($\bar{L}_0$) with $a_0$($\bar{a}_0$).

For the amplitude between the same boundary states $\lambda_1=\lambda_2$, we have the winding mode contribution as
\begin{equation}
\begin{aligned}
\label{eq:Z0}
Z_0 = \sum_{S_1 } g_{\,\!_{\lambda_1} }g_{\,\!_{\lambda_2} } \exp\Big\{- \frac{4\pi}{\beta} 2 \pi g ( \frac{n_1}{ 4 \pi R_1 g} + \frac{m_1 R_1 }{ 2} )^2 \Big\}
\end{aligned}
\end{equation}
In the limit $\beta\rightarrow\infty$, Eq.~\eqref{eq:Z0} can be approximated by a simple two dimensional integral
\begin{equation}\begin{aligned}
Z_0\approx g_{\,\!_{\lambda_1} }g_{\,\!_{\lambda_2} }\beta\int dxdy\exp\Big\{-8 \pi^2 g ( \frac{x}{ 4 \pi R_1 g} + \frac{y R_1 }{ 2} )^2 \Big\}
\end{aligned}\end{equation}
The integral is a constant, thus the winding mode will contribute at most a $\ln\beta$ term to the free energy. Compared to the result in App.~\ref{app:gnd_dn_lambda}-\ref{app:lambda_12}, we conclude that the winding mode contribution is negligible in the large $\beta$ limit.

For the case $\lambda_1\neq\lambda_2$, the partition function will acquire non-zero contribution from the least common divisor of $(n_1,m_1)$ and $(n_2,m_2)$. The same conclusion still holds in the limit $\beta\rightarrow\infty$.

%%% Local Variables:
%%% TeX-master: "bCFT_paper"
%%% TeX-PDF-mode: t
%%% End: