### 2017-05-25 bCFT_paper.tex
[draft] App. G sign issue fixed.

\documentclass[reprint, prb]{revtex4-1}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{bm}
\usepackage[caption=false]{subfig}
\usepackage{todonotes}
\usepackage{comment}
\graphicspath{{./}{./images/}}
\definecolor{darkblue}{rgb}{0.1,0.2,0.6} \definecolor{darkred}{rgb}{0.8,0.1,0.2}
\usepackage[colorlinks,citecolor=darkblue,linkcolor=darkred,urlcolor=darkblue]{hyperref}
\usepackage[all]{hypcap}
\renewcommand{\vec}[1]{\boldsymbol{\mathbf{#1}}}

\newcommand{\cf}{\textit{cf.} } 
\newcommand{\ie}{\textit{i.e.} } 
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\vs}{\textit{vs.} } 
\newcommand{\etal}{\textit{et al.} }
\newcommand{\etc}{\textit{etc.} }

\begin{document}

\title{Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface}

\author{Tianci Zhou}
\email{tzhou13@illinois.edu}
\affiliation{University of Illinois, Department of Physics, 1110 W. Green St. Urbana, IL 61801 USA}

\author{Mao Lin}
\email{maolin2@illinois.edu}
\affiliation{University of Illinois, Department of Physics, 1110 W. Green St. Urbana, IL 61801 USA}

\date{\today}

\begin{abstract}
\input{abstract}
\end{abstract}

\maketitle

\section{Introduction}
\input{intro}

\section{Bosonic Conformal Interface}
\label{sec:bosonic_conformal_interface}
\subsection{General Formulation}
\label{sec_sub:general_formulation}
\input{general_formulation}
\subsection{A Free Boson Lattice Model}
\label{sec_sub:free_boson_lattice}
\input{free_boson_lattice}

\section{Bipartite Fidelity and Loschmidt Echo}
\label{sec:analytic_numerics}
\subsection{Definition}
\input{def_fidel_echo}
\subsection{Notation of Boundary Conditions}
\label{sec:notation}
\input{bd_notation}
\subsection{Analytic Evaluation}
\label{sec_sub:analy_eval}
\input{analy_eval}
\subsection{Numerical Results and Comparison}
\input{num_result}

\section{Discussion}
\label{sec:disc}
\input{disc}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

\begin{acknowledgments}
\todo[inline]{Mao: funding number?}
We are grateful for the discussion and suggestions provided by Thomas Faulkner, Xueda Wen, Shinsei Ryu, Taylor Hughes and Romain Vasseur during the course of this research. 
TZ is supported by the National Science Foundation under grant number NSF-DMR-1306011.
This work made use of the Illinois Campus Cluster, a computing resource that is operated by the Illinois Campus Cluster Program (ICCP) in conjunction with the National Center for Supercomputing Applications (NCSA) and which is supported by funds from the University of Illinois at Urbana-Champaign.
\end{acknowledgments}

\appendix
\section{Corrections to the Free Energy}
\label{app:F_correction}
\input{F_correction}

\section{Conformal Interface in Free Bosonic Lattice}
\label{app:interface_free_boson}
\input{interface_free_boson}

\section{Numerical Computation of Bipartite Fidelity and Loschmidt Echo}
\label{app:comp_fid_echo}
\input{comp_fid_echo}

\section{A Determinant Identity for the Boundary State Amplitude}
\label{app:pf_of_id}
\input{pf_of_id}

\section{Alternative Approach to ${\rm DN} \rightarrow \lambda$ Amplitude}
\label{app:gnd_dn_lambda}
\input{gnd_dn_lambda}

\section{General Boundary State Amplitude}
\label{app:lambda_12}
\input{lambda_1_lambda_2}

\section{Winding Modes of Compact Bosons}
\label{app:compact_diff_boson}
\input{compact_boson}

\bibliographystyle{plain}
\bibliography{bCFT}

\end{document}

%%% Local Variables: 
%%% TeX-PDF-mode: t
%%% End: